Discussion

### 2020-03-17 doc/Discussion.rmd
added refs to discussion

## Within mouse variance

@@

prev reported sex diffferences in

@@

# Evolved Pattern

2.Predictors of Recombination rate evolution within (males)
This section focuses on the largest axis of variation over a short time scale.  We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.

Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength.

## DSBs

Our results find evolution between a subset of strains -- demonstrating that the high rec strains have greater DSB number

Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al). Likely // evidence suggests this pattern is also a feature of the sex differences.

1. More DSBs in the high rec strains

2. The conserved ratios of DSB : CO implies that the increase in DSB numbers are accompanied by proportional increase in crossovers. This is evidence that shifts support away from the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up. 
Together these results indicate that (it's not the CO : NCO decision but an earlier step (when the total number of chromatin loops (the substrate for SPO11, are laid down with the establishment of the meiotic chromosome structure (DNA loop size)

The tethered loop model of DSB formation

3. DSB number being driven by the number of chromatin loops which is positively correlated with axis length.
4. These results supported by the moderate support of longer SC length (pachytene stage in the high rec males).

_the resuts support / replicate those reported in Baier et al_

a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))

-different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures

@@

the correlation of DSB marker with MLH1

(Review Baier et al 2014)

## INTERFERENCE evolution

@@

-- lack of SC length prophase cell charaecterization

-- comparisons within and between sexes (that cattle paper)

>>>>

## Greater Within mouse variance in female

@@

@@

# Evolved Traits

We focused on the largest axis of variation, the faster evolution of genome wide recombination rates in males. One of our primary objectives was to identify other traits which correlate (could explain the increase).

Our results had two notable instances of evolution in the recombination rate (or correlated trait); evolution of precursor number (DSBs) and crossover interference strength. <While SC length and the telomere distances are similar / conserved across the mouse strains>

<We can't distinguish if this is 3 independent instances of evolution in gwRR OR if the patterns are due to shared standing variation, incomplete lineage sorting, but these points are for describing the general patterns.>

# Evolved rates of precursors, DSBs

<define CO:NCO decision, tethered-loop DSB formation SPO11 (cite, paigen? deMassy
setup baier study,  explain >

Our results find evolution between a subset of strains -- demonstrating that the high rec strains have significant higher DSB number.

Our results mirror those found in Baier et al [@baier2014]

Our results show that the ratios of DSB : CO 
**range from X to X which is similar to those presented in [@baier2014] (which report X to X) **

These conserved ratios suggest that the increase in DSB numbers are accompanied by proportional increase in crossovers.

This shifts support away from evolution at the CO:NCO decision to an earlier prophase stage such as when the chromatin compaction is set up. **increase in DSBs is driven by the number of DNA loops** and due to the inverse relationship of loop and chromatin axis length, this would result in longer sc lengths (which reflect the underlying chromatin axis length).

Baier et al found significant differences in sc legnths for 
The tests in our study - show moderate to low significance in the differences in SC lengths. (this is because we have more noise-can't distinguish individual chromosomes)

a study of in house mouse -- using different markers for DSBs and intermediates (markers for the recombination pathway (from DSB to mature crossover))
-different DSB marks, distinct inbred strains two classical lab strains and a wild derived strain from castenus, used FISH for chromosome specific measures

@@

Evolution of recombination rates via the number of precursors -- (has been proposed -- ) (Bairer et al).

# INTERFERENCE evolution

@@

-- lack of SC length prophase cell charaecterization
-- comparisons within and between sexes (that cattle paper)

# Reference
